Hyper-linked Communications: WebRTC enabled asynchronous collaboration

### 2016-05-13 tese2/11_related_work_collaboration.tex
until 1h test case

  \ac{RTP} recorders are independent of payload encoding, they don't decode \ac{RTP} packets, they record packets instead, allowing to record all video and audio formats even if they're encrypted.

@@

\subsection{Media Types}\label{mediatype}~\\
	Media Types can be distinguished by two criteria, the first one describes a media as discrete or continuous, the second one describes it as interactive or non-interactive. A discrete media is characterized by not depending from time, and continuous media as depending on it. Interactive media is characterized by its state being changed by external events such as user interactions.
  %RP estavas a falar de gravar video e passaste a falar de protocolos de streaming. Convém explicar o racional para esta abordagem. Começa a secção a explicar sobre o que vais falar e porquê.
  %RP e usa subsecções...
\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.4\textwidth]{figures/media_types.png}
	\caption{Media Types}
\end{figure}
	For example, an image is non-interactive and discrete, while a video is continuous and non-interactive. A simple collaborative editor with just text is interactive and discrete. An animation that changes in function of user behavior is interactive and continuous.

	Streaming protocols like \ac {RTP} and \ac{RTSP} were designed for continuous and non-interactive media types, such as audio and video. Discrete and non-interactive media don't need to be streamed through \ac{RTP} because they don't change with time. For example, if an image appears in a specific time interval, just the \ac{HTML} or \emph{JavaScript} that will reference the image must be streamed, the image itself is then transferred through \ac{HTTP}.

@@

\begin{table}[!htb]
\centering
\caption{Comparision between Operational Transformation libraries}
\label{table:otcomparision}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Library} & \textbf{Own Server} & \textbf{Own Storage} & \textbf{Operations} \\ \hline
ShareJS          & yes                 & yes                  & text+objects        \\ \hline
TogetherJS       & yes                 & no                   & text+objects        \\ \hline
Goodow           & yes                 & yes                  & text+objects        \\ \hline
Etherpad Lite    & yes                 & yes                  & extendable 			    \\ \hline
OT.js            & no                  & no                   & text                \\ \hline
\end{tabular}
\end{table}

>>>>

  \ac{RTP} recorders are independent of payload encoding, they do not decode \ac{RTP} packets, they record packets instead, allowing to record all video and audio formats even if they're encrypted.

@@

\subsection{Media Types}\label{mediatype}~\\
	Media Types can be distinguished by two criteria, the first one describes a media as discrete or continuous, the second one describes it as interactive or non-interactive. A discrete media is characterized by not depending from time, and continuous media as depending on it. Interactive media is characterized by its state being changed by external events such as user interactions.
  %RP estavas a falar de gravar video e passaste a falar de protocolos de streaming. Convém explicar o racional para esta abordagem. Começa a secção a explicar sobre o que vais falar e porquê.
  %RP e usa subsecções...
\begin{figure}
	\centering
	\includegraphics[width=0.4\textwidth]{figures/media_types.png}
	\caption{Media Types}
\end{figure}
	For example, an image is non-interactive and discrete, while a video is continuous and non-interactive. A simple collaborative editor with just text is interactive and discrete. An animation that changes in function of user behavior is interactive and continuous.

	Streaming protocols like \ac {RTP} and \ac{RTSP} were designed for continuous and non-interactive media types, such as audio and video. Discrete and non-interactive media do not need to be streamed through \ac{RTP} because they do not change with time. For example, if an image appears in a specific time interval, just the \ac{HTML} or \emph{JavaScript} that will reference the image must be streamed, the image itself is then transferred through \ac{HTTP}.

@@

\begin{table}
\centering
\caption{Comparision between Operational Transformation libraries}
\label{table:otcomparision}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Library} & \textbf{Own Server} & \textbf{Own Storage} & \textbf{Operations} \\ \hline
ShareJS          & yes                 & yes                  & text+objects        \\ \hline
TogetherJS       & yes                 & no                   & text+objects        \\ \hline
Goodow           & yes                 & yes                  & text+objects        \\ \hline
Etherpad Lite    & yes                 & yes                  & extendable 			    \\ \hline
OT.js            & no                  & no                   & text                \\ \hline
\end{tabular}
\end{table}

### 2016-05-13 tese2/13_implementation_model.tex
until 1h test case

    Each \emph{MongoDB} document is represented by a \emph{JSON} object and, as a result, each document may have different attributes within a collection. Needless to say that, by following this approach we don't need to create the collections with a well predefined schema. In fact we don't need to define it at all but, for reasons of coherence and organization, we represent our database collections as they would have a predefined schema by following the same document structure as we will present in this section.
    %RP a última frase é confusa. O que queres dizer? Reescreve
    %HR está melhor?

@@

\begin{table}[!htb]
\centering
\caption{Schema representation}
\label{table:schema}
    \begin{tabular}{|ll|}
        \hline
        \multicolumn{2}{|c|}{\textbf{Collection name}}            \\ \hline
        $\Diamondblack$ \underline{\_id (primary key)} & ObjectId \\ 
        $\medbullet$ Not nullable property name & Property type   \\ 
        $\medcirc$ Nullable property name       & Property type   \\
        $\medcirc$ Reference to document        & ObjectId        \\
        $\medbullet$ Embedded document          & Document        \\ 
        $\medbullet$ Embedded list              & List[Type]      \\ \hline
    \end{tabular}
\end{table}

@@

\begin{table}[!htb]
\centering
\caption{Generic data model}
\label{table:generic}
    \begin{tabular}{cc}
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{Data}}          \\ \hline
            $\Diamondblack$ \underline{\_id} & ObjectId  \\ 
            $\medbullet$ owner          & ObjectId       \\ 
            $\medbullet$ properties    & List[Attribute] \\ 
            $\medcirc$ searchableValues & List[Text]     \\ \hline
        \end{tabular}
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{Attribute}}     \\ \hline
            $\medbullet$ key            & Text           \\ 
            $\medcirc$ value            & Object         \\ 
            $\medbullet$ identifiable   & Boolean        \\ 
            $\medcirc$ readPermissions  & List[ObjectId] \\ 
            $\medcirc$ writePermissions & List[ObjectId] \\ \hline
        \end{tabular}    
    \end{tabular}
\end{table}

@@

Another important attribute specification is the owner identifiability, which tells us if the attribute identifies the object. This specification lets us create abstract authentication services. For example a user can login into our system by providing any attribute that identifies himself, e.g. the e-mail but others are possible like the user name or cellphone number.

@@

\begin{table}[!htb]
\centering
\caption{User model}
\label{table:user}
    \begin{tabular}{|ll|}
        \hline
        \multicolumn{2}{|c|}{\textbf{User}}         \\ \hline
        $\Diamondblack$ \underline{\_id}  & ObjectId  \\ 
        $\medbullet$ hash           & Text          \\ 
        $\medbullet$ salt      & Text               \\ \hline
    \end{tabular}
\end{table}

@@

\begin{table}[!htb]
\centering
\caption{Relation model}
\label{table:relation}
    \begin{tabular}{|ll|}
        \hline
        \multicolumn{2}{|c|}{\textbf{Relation}}     \\ \hline
        $\Diamondblack$ \underline{\_id}  & ObjectId  \\ 
        $\medbullet$ source           & ObjectId    \\ 
        $\medbullet$ target      & ObjectId         \\ \hline
    \end{tabular}
\end{table}

@@

\begin{table}[!htb]
\centering
\caption{Group model}
\label{table:group}
    \begin{tabular}{cc}
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{Group}}        \\ \hline
            $\Diamondblack$ \underline{\_id}  & ObjectId  \\ 
            $\medcirc$ inviteToken           & Text     \\ 
            $\medbullet$ visibility      & Text         \\ \hline
        \end{tabular}

@@

\begin{table}[!htb]
\centering
\caption{Message model}
\label{table:message}
    \begin{tabular}{|ll|}
        \hline
        \multicolumn{2}{|c|}{\textbf{Message}}      \\ \hline
        $\Diamondblack$ \underline{\_id}  & ObjectId  \\ 
        $\medbullet$ source           & ObjectId    \\ 
        $\medbullet$ target      & ObjectId         \\ 
        $\medbullet$ content      & Text            \\ \hline
    \end{tabular}
\end{table}

@@

\begin{table}[!htb]
\centering
\caption{Hyper content model}
\label{table:hyper}
    \begin{tabular}{cc}
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{HyperContent}}  \\ \hline
            $\Diamondblack$ \underline{\_id}  & ObjectId \\ 
            $\medbullet$ groupId            & ObjectId   \\ 
            $\medbullet$ start              & Date       \\ 
            $\medbullet$ end                & Date       \\ 
            $\medbullet$ content            & Text       \\ 
            $\medbullet$ searchableContent  & Text       \\ \hline
        \end{tabular}

@@

\begin{table}[!htb]
\centering
    \caption{Collaborative content model}
    \label{table:collaborative}
    \begin{tabular}{|ll|}
        \hline
        \multicolumn{2}{|c|}{\textbf{CollaborativeContent}} \\ \hline
        $\Diamondblack$ \underline{\_id}  & ObjectId        \\ 
        $\medbullet$ groupId            & ObjectId          \\ 
        $\medcirc$ content            & Text                \\ \hline
    \end{tabular}
\end{table}

@@

\begin{table}[!htb]
\centering
    \caption{Recording model}
    \label{table:recording}
    \begin{tabular}{cc}
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{RecordingInterval}}  \\ \hline
            $\Diamondblack$ \underline{\_id}  & ObjectId      \\ 
            $\medbullet$ groupId            & ObjectId        \\ 
             $\medbullet$ start             & Date            \\ 
            $\medbullet$ end                & Date            \\ \hline
        \end{tabular}   
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{RecordingChunk}}  \\ \hline
            $\Diamondblack$ \underline{\_id}  & ObjectId     \\ 
            $\medbullet$ groupId            & ObjectId     \\ 
            $\medbullet$ owner           & ObjectId        \\ 
            $\medbullet$ sessionId          & Text         \\
            $\medbullet$ start              & Date         \\ 
            $\medbullet$ end                & Date         \\ 
            $\medbullet$ url               & Text          \\ \hline
        \end{tabular}
    \end{tabular}
\end{table}

>>>>

    Each \emph{MongoDB} document is represented by a \emph{JSON} object and, as a result, each document may have different attributes within a collection. Needless to say that, by following this approach we do not need to create the collections with a well predefined schema. In fact we do not need to define it at all but, for reasons of coherence and organization, we represent our database collections as they would have a predefined schema by following the same document structure as we will present in this section.
    %RP a última frase é confusa. O que queres dizer? Reescreve
    %HR está melhor?

@@

\begin{table}
\centering
\caption{Schema representation}
\label{table:schema}
    \begin{tabular}{|ll|}
        \hline
        \multicolumn{2}{|c|}{\textbf{Collection name}}            \\ \hline
        $\Diamondblack$ \underline{\_id (primary key)} & ObjectId \\ 
        $\medbullet$ Not nullable property name & Property type   \\ 
        $\medcirc$ Nullable property name       & Property type   \\
        $\medcirc$ Reference to document        & ObjectId        \\
        $\medbullet$ Embedded document          & Document        \\ 
        $\medbullet$ Embedded list              & List[Type]      \\ \hline
    \end{tabular}
\end{table}

@@

\begin{table}
\centering
\caption{Generic data model}
\label{table:generic}
    \begin{tabular}{cc}
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{Data}}          \\ \hline
            $\Diamondblack$ \underline{\_id} & ObjectId  \\ 
            $\medbullet$ owner          & ObjectId       \\ 
            $\medbullet$ properties    & List[Attribute] \\ 
            $\medcirc$ searchableValues & List[Text]     \\ \hline
        \end{tabular}
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{Attribute}}     \\ \hline
            $\medbullet$ key            & Text           \\ 
            $\medcirc$ value            & Object         \\ 
            $\medbullet$ identifiable   & Boolean        \\ 
            $\medcirc$ readPermissions  & List[ObjectId] \\ 
            $\medcirc$ writePermissions & List[ObjectId] \\ \hline
        \end{tabular}    
    \end{tabular}
\end{table}

@@

Another important attribute specification is the owner identifiability, which tells us if the attribute identifies the object. This specification lets us create abstract authentication services. For example a user can login into our system by providing any attribute that identifies himself, \emph{e.g.} the e-mail but others are possible like the user name or cellphone number.

@@

\begin{table}
\centering
\caption{User model}
\label{table:user}
    \begin{tabular}{|ll|}
        \hline
        \multicolumn{2}{|c|}{\textbf{User}}         \\ \hline
        $\Diamondblack$ \underline{\_id}  & ObjectId  \\ 
        $\medbullet$ hash           & Text          \\ 
        $\medbullet$ salt      & Text               \\ \hline
    \end{tabular}
\end{table}

@@

\begin{table}
\centering
\caption{Relation model}
\label{table:relation}
    \begin{tabular}{|ll|}
        \hline
        \multicolumn{2}{|c|}{\textbf{Relation}}     \\ \hline
        $\Diamondblack$ \underline{\_id}  & ObjectId  \\ 
        $\medbullet$ source           & ObjectId    \\ 
        $\medbullet$ target      & ObjectId         \\ \hline
    \end{tabular}
\end{table}

@@

\begin{table}
\centering
\caption{Group model}
\label{table:group}
    \begin{tabular}{cc}
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{Group}}        \\ \hline
            $\Diamondblack$ \underline{\_id}  & ObjectId  \\ 
            $\medcirc$ inviteToken           & Text     \\ 
            $\medbullet$ visibility      & Text         \\ \hline
        \end{tabular}

@@

\begin{table}
\centering
\caption{Message model}
\label{table:message}
    \begin{tabular}{|ll|}
        \hline
        \multicolumn{2}{|c|}{\textbf{Message}}      \\ \hline
        $\Diamondblack$ \underline{\_id}  & ObjectId  \\ 
        $\medbullet$ source           & ObjectId    \\ 
        $\medbullet$ target      & ObjectId         \\ 
        $\medbullet$ content      & Text            \\ \hline
    \end{tabular}
\end{table}

@@

\begin{table}
\centering
\caption{Hyper content model}
\label{table:hyper}
    \begin{tabular}{cc}
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{HyperContent}}  \\ \hline
            $\Diamondblack$ \underline{\_id}  & ObjectId \\ 
            $\medbullet$ groupId            & ObjectId   \\ 
            $\medbullet$ start              & Date       \\ 
            $\medbullet$ end                & Date       \\ 
            $\medbullet$ content            & Text       \\ 
            $\medbullet$ searchableContent  & Text       \\ \hline
        \end{tabular}

@@

\begin{table}
\centering
    \caption{Collaborative content model}
    \label{table:collaborative}
    \begin{tabular}{|ll|}
        \hline
        \multicolumn{2}{|c|}{\textbf{CollaborativeContent}} \\ \hline
        $\Diamondblack$ \underline{\_id}  & ObjectId        \\ 
        $\medbullet$ groupId            & ObjectId          \\ 
        $\medcirc$ content            & Text                \\ \hline
    \end{tabular}
\end{table}

@@

\begin{table}
\centering
    \caption{Recording model}
    \label{table:recording}
    \begin{tabular}{cc}
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{RecordingInterval}}  \\ \hline
            $\Diamondblack$ \underline{\_id}  & ObjectId      \\ 
            $\medbullet$ groupId            & ObjectId        \\ 
             $\medbullet$ start             & Date            \\ 
            $\medbullet$ end                & Date            \\ \hline
        \end{tabular}   
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{RecordingChunk}}  \\ \hline
            $\Diamondblack$ \underline{\_id}  & ObjectId     \\ 
            $\medbullet$ groupId            & ObjectId     \\ 
            $\medbullet$ owner           & ObjectId        \\ 
            $\medbullet$ sessionId          & Text         \\
            $\medbullet$ start              & Date         \\ 
            $\medbullet$ end                & Date         \\ 
            $\medbullet$ url               & Text          \\ \hline
        \end{tabular}
    \end{tabular}
\end{table}

### 2016-05-13 tese2/14_implementation_signaling.tex
until 1h test case

\begin{figure}[!htb]
    \centering
    \begin{subfigure}{}
    	\includegraphics[width=0.9\textwidth]{figures/signaling}
    \end{subfigure}
    \caption{Signaling sequence diagram}
    \label{fig:signaling2}
\end{figure}

@@

Instead, if we create a \emph{WebSocket} \ac{API}, we only need one \ac{TCP} connection and, at the same, provide bi-directional communications without additional headers. Moreover, using \emph{WebSocket} allows the application server to send messages asynchronously to the client without the client having to request it, which can be slightly faster than using \ac{HTTP} which is based on a request followed by a response.
%RP em vez de slighhtly faster não queres antes enfatizar que pode ser antes o servidor a iniciar a comunicação, de forma assincrona?
%HR done
Our signaling protocol consists of sending and receiving \ac{JSON} formated messages, e.g. listing \ref{lst:msgtype}, over WebSockets by both the application server and the client.

@@

At this stage, the web application's user is asked if he wants to share its camera and microphone, share screen or just receive streams from the server. If the user decides to perform a screen share, \emph{adapter.js}\footnote{\url{https://github.com/Temasys/AdapterJS}(accessed March 15, 2016).} may ask to install a plug-in if the browser doesn't support screen sharing \footnote{\url{http://iswebrtcreadyyet.com/} (Accessed May 11, 2016)}.
%RP web app user? OU apenas client user?
%RP foot note com exemplos de browsers que aceitem screen sharing?
%HR done
If the user decides to share either from camera or screen, \emph{getUserMedia} is called with the correspondent constraints in order to obtain a local stream. We use the constraints presented on listing \ref{lst:constraints}.

@@

The \emph{local session description} contains the session identifier, codecs, containers, transport protocols and ports used per media type. The \emph{local session description} is useful to conclude if the client is receiving only, which means that \ac{KMS} doesn't need to mix, record nor analyze streams coming from the user's \ac{WebRTC} endpoint.

>>>>

\begin{figure}
    \centering
    \begin{subfigure}{}
    	\includegraphics[width=0.9\textwidth]{figures/signaling}
    \end{subfigure}
    \caption{Signaling sequence diagram}
    \label{fig:signaling2}
\end{figure}

@@

Instead, if we create a \emph{WebSocket} \ac{API}, we only need one \ac{TCP} connection and, at the same, provide bi-directional communications without additional headers. Moreover, using \emph{WebSocket} allows the application server to send messages asynchronously to the client without the client having to request it, which can be slightly faster than using \ac{HTTP} which is based on a request followed by a response.
%RP em vez de slighhtly faster não queres antes enfatizar que pode ser antes o servidor a iniciar a comunicação, de forma assincrona?
%HR done
Our signaling protocol consists of sending and receiving \ac{JSON} formated messages, \emph{e.g.} listing \ref{lst:msgtype}, over WebSockets by both the application server and the client.

@@

At this stage, the web application's user is asked if he wants to share its camera and microphone, share screen or just receive streams from the server. If the user decides to perform a screen share, \emph{adapter.js}\footnote{\url{https://github.com/Temasys/AdapterJS}(accessed March 15, 2016).} may ask to install a plug-in if the browser does not support screen sharing \footnote{\url{http://iswebrtcreadyyet.com/} (Accessed May 11, 2016)}.
%RP web app user? OU apenas client user?
%RP foot note com exemplos de browsers que aceitem screen sharing?
%HR done
If the user decides to share either from camera or screen, \emph{getUserMedia} is called with the correspondent constraints in order to obtain a local stream. We use the constraints presented on listing \ref{lst:constraints}.

@@

The \emph{local session description} contains the session identifier, codecs, containers, transport protocols and ports used per media type. The \emph{local session description} is useful to conclude if the client is receiving only, which means that \ac{KMS} does not need to mix, record nor analyze streams coming from the user's \ac{WebRTC} endpoint.

### 2016-05-13 tese2/15_implementation_recording.tex

	The process to play a video was quite simple. The user specified which user and date he intended to play, the server calculated the intersection between the requested date and the block bounds and returned the file to the client.
	Although this idea was fairly simple, we couldn't achieve seamless sequential block switching. Downloading the video file always took a noticeable amount of time. To solve this problem, while we were playing a block, the next one was downloading in parallel, so when the current block finished playing we would have an available block to play.

@@

	One of the features that \emph{Kurento Repository} provides is the ability to play directly from the database without having to download the entire file to \ac{KMS}. The same is true for recording, but because the file headers are in the beginning and the file is written until is stops, the headers don't contain the necessary information for seeking the file.
	Although we gain with scalability with this approach, we lose access over the file for changing it to fit our needs, namely for using \emph{ffmpeg} or other video manipulation tool.

>>>>

	The process to play a video was quite simple. The user specified which user and date he intended to play, the server calculated the intersection between the requested date and the block bounds and returned the file to the client.
	Although this idea was fairly simple, we could not achieve seamless sequential block switching. Downloading the video file always took a noticeable amount of time. To solve this problem, while we were playing a block, the next one was downloading in parallel, so when the current block finished playing we would have an available block to play.

@@

	One of the features that \emph{Kurento Repository} provides is the ability to play directly from the database without having to download the entire file to \ac{KMS}. The same is true for recording, but because the file headers are in the beginning and the file is written until is stops, the headers do not contain the necessary information for seeking the file.
	Although we gain with scalability with this approach, we lose access over the file for changing it to fit our needs, namely for using \emph{ffmpeg} or other video manipulation tool.

### 2016-05-13 tese2/16_implementation_content.tex

	As manually content insertion is a laborious task that can be realized after the video is recorder, we provide an alternative mechanism for real time introduction of superimposed content.	In order to help the content creator to introduce and synchronize its content in real-time, we allow the user to encode its content into \emph{QR codes} and show it to the camera in real time.

	\ac{KMS} lets registering event handlers for \emph{QR code} detection. The component that detects \emph{QR codes} on \ac{KMS} is called periodically and fires the handlers with the decoded content. This mechanism does not detect if the \emph{QR code} enters or leaves the screen. We had to implement our own mechanism for detecting those events. Each user session in the server maintains a map with the contents that are present on the screen. In order to apply our algorithm more efficiently we calculate the content's hash through the \emph{md5} method. If the hash is not present in the map, it means the \emph{QR code} was entering the screen, we add that hash to the map and associate the current time to it, all the users are notified to watch that content in real-time. If the same \emph{QR code} is not detected after a time bigger then two periods we can conclude that the \emph{QR code} left the screen and we add the correspondent content into our database. Listing \ref{lst:pseudo_qrcode} shows the pseudo code for \emph{QR code} leaving detection.

>>>>

	As manually content insertion is a laborious task that can be realized after the video is recorder, we provide an alternative mechanism for real time introduction of superimposed content.	In order to help the content creator to introduce and synchronize its content in real-time, we allow the user to encode its content into \ac{QR} codes and show it to the camera in real time.

	\ac{KMS} lets registering event handlers for \ac{QR} code detection. The component that detects \ac{QR} codes on \ac{KMS} is called periodically and fires the handlers with the decoded content. This mechanism does not detect if the \ac{QR} code enters or leaves the screen. We had to implement our own mechanism for detecting those events. Each user session in the server maintains a map with the contents that are present on the screen. In order to apply our algorithm more efficiently we calculate the content's hash through the \emph{md5} method. If the hash is not present in the map, it means the \ac{QR} code was entering the screen, we add that hash to the map and associate the current time to it, all the users are notified to watch that content in real-time. If the same \ac{QR} code is not detected after a time bigger then two periods we can conclude that the \ac{QR} code left the screen and we add the correspondent content into our database. Listing \ref{lst:pseudo_qrcode} shows the pseudo code for \ac{QR} code leaving detection.

### 2016-05-13 tese2/acknowledgements.tex

I would like to thank my dissertation supervisor Ricardo Pereira and co-supervisor Paulo Chainho for their active support and ideas that were crucial for the development of this thesis.

@@

Not less important I should thank all IST employees that maintained the space I used during this journey.

>>>>

I would like to thank my thesis supervisor Ricardo Pereira and co-supervisor Paulo Chainho for their active support and ideas that were crucial for the development of this thesis.

@@

Not less important, I should thank all IST employees that maintained the space I used during this journey.

### 2016-05-13 tese2/evaluation.tex

  Just for testing purposes, we performed the same performance tests disabling \emph{QR codes} detection in order to understand how \ac{CPU} intensive this task is. Figure \ref{fig:test_without_qrcode_cpu} shows the results for the same experience but with \emph{QR code} detection disabled.

@@

We conclude that \emph{QR code} detection is a very intensive task, approximately doubling the amount of work performed by the \ac{CPU}s. Without this feature, the network and memory usage had insignificant changes compared to \ac{CPU} usage.

@@

      \item Enter a conference room by sharing their camera and add a subtitle in the video, preview it, specify an interval of time and save it. Then show the provided \emph{QR code} to the camera and leave the leave the conference room.
      \end{enumerate}

@@

On the fourth task the users didn't placed their annotations on right place. We only provide one way to place annotations, which is by dragging them on the time-line. 
After creating an annotation, users also didn't understand that they had to save their changes on the time-line. We solved this problem by saving automatically each time the user changed any annotation. The old user interface can be seen on figure \ref{fig:test_ui_05}.

@@

The three users had success with the fifth task, but they revealed difficulties choosing the time interval for the hyper content. We are aware that this is a difficult task and that's why we have introduced the \emph{QR code} content creation mechanism which was very easy to learn (with a \emph{QR code} already printed into paper).

Although content creation through \emph{QR codes} revealed much easier to perform, this approach implies showing and, as a consequence, recording video with the \emph{QR codes}, making them visible to other users. In order to solve this problem, we could add a new video sharing mode that allow users to send video to \ac{KMS} without composing, recording and showing to other users.

>>>>

  Just for testing purposes, we performed the same performance tests disabling \ac{QR} codes detection in order to understand how \ac{CPU} intensive this task is. Figure \ref{fig:test_without_qrcode_cpu} shows the results for the same experience but with \ac{QR} code detection disabled.

@@

We conclude that \ac{QR} code detection is a very intensive task, approximately doubling the amount of work performed by the \ac{CPU}s. Without this feature, the network and memory usage had insignificant changes compared to \ac{CPU} usage.

@@

      \item Enter a conference room by sharing their camera and add a subtitle in the video, preview it, specify an interval of time and save it. Then show the provided \ac{QR} code to the camera and leave the leave the conference room.
      \end{enumerate}

@@

On the fourth task the users did not placed their annotations on right place. We only provide one way to place annotations, which is by dragging them on the time-line. 
After creating an annotation, users also did not understand that they had to save their changes on the time-line. We solved this problem by saving automatically each time the user changed any annotation. The old user interface can be seen on figure \ref{fig:test_ui_05}.

@@

The three users had success with the fifth task, but they revealed difficulties choosing the time interval for the hyper content. We are aware that this is a difficult task and that's why we have introduced the \ac{QR} code content creation mechanism which was very easy to learn (with a \ac{QR} code already printed into paper).

Although content creation through \ac{QR} codes revealed much easier to perform, this approach implies showing and, as a consequence, recording video with the \ac{QR} codes, making them visible to other users. In order to solve this problem, we could add a new video sharing mode that allow users to send video to \ac{KMS} without composing, recording and showing to other users.

### 2016-05-13 tese2/frontCover.tex

\begin{center}
%
% Figure (Image or plot)
\vspace{0.3cm}
% TODO - Use of an image is optional
\hspace{-0.5cm}\includegraphics[height=50mm]{Figures/logo.png}
% TODO - If you don't use an image uncomment the next line
% \vspace{5.0cm}

% Title, author and degree
\vspace{0.8cm}
{\FontLb Hyper-linked Communications: WebRTC enabled asynchronous collaboration} \\
%\vspace{0.2cm}
%{\FontMn Subtítulo (facultativo)} \\
%\vspace{1.9cm}
\vspace{2.6cm}
{\FontMb Henrique Duarte Lopes Rocha} \\
\vspace{1.9cm}
{\FontLn Disserta\c{c}\~{a}o para a obten\c{c}\~{a}o de Grau de Mestre em} \\
\vspace{0.3cm}
{\FontLb Engenharia de Telecomunicações e Informática} \\
\vspace{1.9cm}
{\FontMb J\'{u}ri} \\
\vspace{0.3cm}
{\FontSn %
Presidente:        Nome do Presidente \\
Orientador:        Nome do Orientador \\
Co-orientador:     Nome do Co-orientador \\
Vogais:            Nome do Vogal 1 \\
                   Nome do Vogal 2 \\
}
\vspace{1.5cm}

{\FontMb M\^{e}s e Ano} \\
%
\end{center}

>>>>

\begin{center}
%
% Figure (Image or plot)
\vspace{0.3cm}
% TODO - Use of an image is optional
\hspace{-0.5cm}\includegraphics[height=50mm]{Figures/logo.png}
% TODO - If you do not use an image uncomment the next line
% \vspace{5.0cm}

% Title, author and degree
\vspace{0.8cm}
{\FontLb Hyper-linked Communications: WebRTC enabled asynchronous collaboration} \\
%\vspace{0.2cm}
%{\FontMn Subtítulo (facultativo)} \\
%\vspace{1.9cm}
\vspace{2.6cm}
{\FontMb Henrique Duarte Lopes Rocha} \\
\vspace{1.9cm}
{\FontLn Dissertation submitted to obtain the Master Degree in} \\
\vspace{0.3cm}
{\FontLb Communications Networks Engineering} \\
\vspace{1.9cm}
{\FontMb J\'{u}ri} \\
\vspace{0.3cm}
{\FontSn %
Chairperson: 	   Prof. Paulo Jorge Pires Ferreira
Supervisor:        Prof. Ricardo Jorge Feliciano Lopes Pereira \\
Members of the Committee: Prof. João Pedro Barreto
}
\vspace{1.5cm}

{\FontMb May 2016} \\
%
\end{center}

### 2016-05-13 tese2/implementation.tex

As a consequence of not using \ac{XMPP} for the signaling protocol, we have defined our own protocol based on \emph{WebSockets} for performing the stream between clients and \ac{KMS}. From the \ac{KMS} functionalities, our solution uses the recording feature, composite endpoints that mix multiple audio and video streams into a single one and \emph{QR code} detection for creating hyper content.

>>>>

As a consequence of not using \ac{XMPP} for the signaling protocol, we have defined our own protocol based on \emph{WebSockets} for performing the stream between clients and \ac{KMS}. From the \ac{KMS} functionalities, our solution uses the recording feature, composite endpoints that mix multiple audio and video streams into a single one and \ac{QR} code detection for creating hyper content.

### 2016-05-13 tese2/introduction.tex

@@

	As communications technologies appeared, we adapted the way we communicate. {\color{blue} The purpose of this project is not the replacement of} the current video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.

	{\color{blue}With the advent of WebRTC and its successive integration with web browsers, it became possible to develop video conference web applications without plugins, this presents a range of possibilities on what can be implemented using already existing web technologies.}

	For multiple reasons, we often need to repeat or postpone some of our tasks, some people tend to forget what they ear or see.
        %RP Parece-me que até aqui ainda é background ou motivação. Não me parece que estejas já a descrever a solução proposta.

@@

	{\color{blue}Another point to consider is making available collaborative tools for text edition, instant messaging and video superimposed content creation and visualization.}

@@

Making it clear, this project aims to complement current audio, text and video communications in order to create rich and collaborative interfaces with the ability to add more content on a future time (e.g. creating time annotations for improving content search) in order to increase its value. It is also important to highlight another goal of this project which is the ability to navigate in time by rewinding communications, fast-forward and jump to certain points.

@@

A web application with an easy to learn user interface will be developed to accomplish solving our problem. Our application, unnamed yet, is targeted at web browsers that are compatible with only standard technologies like JavaScript, \ac{WebRTC}, \ac{HTML}5 and \ac{CSS}3. Any additional plug-in is avoidable, \emph{JavaScript} libraries will be preferred as they can be downloaded on the fly.

>>>>

	As communications technologies appeared, we adapted the way we communicate. The purpose of this project is not the replacement of the current video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.

	With the advent of WebRTC and its successive integration with web browsers, it became possible to develop video conference web applications without plugins. This presents a range of possibilities on what can be implemented using already existing web technologies.

        %RP Parece-me que até aqui ainda é background ou motivação. Não me parece que estejas já a descrever a solução proposta.

@@

	For multiple reasons, we often need to repeat or postpone some of our tasks, some people tend to forget what they hear or see.

@@

	Our goal in this project is to develop an application targeted to the web platform, resorting to \ac{WebRTC}, that leverages the hyper-linked communications by providing a video conference environment enriched with interactive and non-interactive discrete media types such as images, subtitles, forms and all types of content that can be added using \ac{HTML}5, \ac{CSS}3 and \emph{JavaScript} including continuous media types such as video, music and animations.

	One of the key features of this project is the ability to navigate in time in order to reproduce the conversation again or introduce hyper-content to it such as time annotations, interactive lists of topics and subtitles. In this context we also provide a simpler method for creating and synchronizing hyper-content using \ac{QR} codes.

	In addition to this conference environment, which provides different functionalities than traditional conference environments such as \emph{Skype} and \emph{Google Hangouts}, we also enable a collaborative text editor and a chat that supports sending time hyper-links and files to conference participants.

	Furthermore, another relevant feature is the possibility to compose multiple video streams into a single one, which enables adding more users to conference rooms without impacting on clients performance. Users can change to individual streams on demand or automatically to the talking users.

@@

Making it clear, this project aims to complement current audio, text and video communications in order to create rich and collaborative interfaces with the ability to add more content on a future time (\emph{e.g.} creating time annotations for improving content search) in order to increase its value. It is also important to highlight another goal of this project which is the ability to navigate in time by rewinding communications, fast-forward and jump to certain points.

@@

A web application with an easy to learn user interface was developed to accomplish solving our problem. Our application, unnamed yet, is targeted at web browsers that are compatible with only standard technologies like JavaScript, \ac{WebRTC}, \ac{HTML}5 and \ac{CSS}3. Any additional plug-in was avoided, \emph{JavaScript} libraries were preferred as they can be downloaded on-the-fly.
